Kindly Bent to Free Us

### 2019-07-03 kindly/infer/infer.tex
Fix weakening.
Progress on inference soundness proof.

We now formulate type inference for the \lang language. Type inference is
based on the HM(X) framework~\citep{DBLP:journals/tapos/OderskySW99} which
presents how to implement Hindley-Milner type inference in a language
with qualified types. We first present various preliminaries definitions,
we then show our extension of HM(X) that support a kind system and affine
types.

@@

\begin{figure}[tp]\centering
\begin{mathpar}
  % \Sv\cap\Sv'
  % &= \left\{ x \to l,l'
  %   \mid \Sone{x}{l}\in\Sv \text{ and } \Sone{x}{l'}\in\Sv' \right\} \\
  % \infer
  % { \forall x \in \Dom{\Sv_1}\cup\Dom{\Sv_2}.\quad
  %   \bsplit{C_x}{\Sv(x)}{\Sv_1(x)}{\Sv_2(x)}
  % } 
  % { \bsplit{(\bigwedge C_x)}{\Sv}{\Sv_1}{\Sv_2} }
  % &= \Sv\cap\Sv' \sqcup
  %   (\Sv\Sdel{x' \mid x'\in\Sv'}) \sqcup
  %   (\Sv'\Sdel{x \mid x\in\Sv})
  %   &\text{where $\sqcup$ is the disjoint union}\\
  % \and
  \inferrule{}
  { \Cleq{\Sv}{k}
    = \bigwedge_{\Sone{x}{\schm}\in\Sv} \Cleq{\schm}{k}
  }
  \and
  % \inferrule
  % { \forall \bshadow{x} \in \Sv, b'\in\Sv'.\quad
  %   \tregion{C_x}{y}{\bshadow{x}}{b'} }
  % { \bregion{(\bigwedge C_x)}{y}{\Sv}{\Sv'} }
  % \and
  % \inferrule
  % { \text{if } \operatorname{lenght}(\Sv(x))  = \Cempty{} \wedge  }
  { \Weaken_{(x,k)}(\Sv) =
    \begin{dcases}
      \Ctrue& \operatorname{length}(\Sv(x)) = 1\\
      {k}\lk\kaff_\infty& \text{ otherwise}\\
    \end{dcases}

  }
\end{mathpar}
\caption{Operation on usage maps $\Sv$}
\label{op:usgmap}
\end{figure}

@@

We note $\inferK{(C,\unif)}{\bf{\E}}{\bf{\tau}}{k}$ when type $\tau$ has kind $k$
in environment $\E$ under constraints $C$ and unifier $\unif$. From an
algorithmic point of view, $\E$ and $\tau$ are the input parameters of
our inference procedure.
We present the kind inference algorithm as a set of syntax-directed rules in
\cref{rules:kinding}.
Since higher-kinded types are not supported, the type application
rule and the type constructor rule are merged in {\sc KApp}.
Additionally, type variables must be of a simple kind in rule {\sc KVar}.
Note that in the case of a type constructor with no argument, the {\sc KApp}
rule degenerates to a simpler form which is similar to the {\sc KVar} rule.
Kind schemes are instantiated in the {\sc KApp} and {\sc KVar} rules by creating
fresh kind variables and the associated substitution.

>>>>

We now formulate type inference for the \lang language. Type inference is
based on the HM(X) framework~\citep{DBLP:journals/tapos/OderskySW99} which
presents how to implement Hindley-Milner type inference in a language
with qualified types and an arbitrary custom constraint language.
We first present various preliminaries definitions,
we then show our extension of HM(X) that support a kind system and affine
types.

@@

\begin{figure}[tp]\centering
\begin{mathpar}
  % \Sv\cap\Sv'
  % &= \left\{ x \to l,l'
  %   \mid \Sone{x}{l}\in\Sv \text{ and } \Sone{x}{l'}\in\Sv' \right\} \\
  % \infer
  % { \forall x \in \Dom{\Sv_1}\cup\Dom{\Sv_2}.\quad
  %   \bsplit{C_x}{\Sv(x)}{\Sv_1(x)}{\Sv_2(x)}
  % } 
  % { \bsplit{(\bigwedge C_x)}{\Sv}{\Sv_1}{\Sv_2} }
  % &= \Sv\cap\Sv' \sqcup
  %   (\Sv\Sdel{x' \mid x'\in\Sv'}) \sqcup
  %   (\Sv'\Sdel{x \mid x\in\Sv})
  %   &\text{where $\sqcup$ is the disjoint union}\\
  % \and
  % \inferrule{}
  % { \Cleq{\Sv}{k}
  %   = \bigwedge_{\Sone{x}{\schm}\in\Sv} \Cleq{\schm}{k}
  % }
  % \and
  % \inferrule
  % { \forall \bshadow{x} \in \Sv, b'\in\Sv'.\quad
  %   \tregion{C_x}{y}{\bshadow{x}}{b'} }
  % { \bregion{(\bigwedge C_x)}{y}{\Sv}{\Sv'} }
  % \and
  % \inferrule
  % { \text{if } \operatorname{lenght}(\Sv(x))  = \Cempty{} \wedge  }
  { \Weaken_{\bvar{x}{\sigma}}(\Sv) =
    \begin{dcases}
      \Ctrue& \text{if } x\in\Sv\\
      \Cleq{\sigma}{\kaff_\infty}&\\
    \end{dcases}

  }

  \inferrule{
    \inferK{C \Cand C_x} \E \tau {k'} \\
    D = C \Cand C_x \Cand \Cleq{k'}k
  }{
    \E \vdash \Cleq{\forall \kvar_i \forall (\tvar_j:k_j).\ \qual{C_x}{\tau}}{k} \Crewrite  D
  }
\end{mathpar}
\caption{Auxiliary rules for inference}
\label{op:usgmap}
\end{figure}

@@

We note $\inferK{(C,\unif)}{\bf{\E}}{\bf{\tau}}{k}$ when type $\tau$ has kind $k$
in environment $\E$ under constraints $C$ and unifier $\unif$. From an
algorithmic point of view, $\E$ and $\tau$ are the input parameters of
our inference procedure.
We present the kind inference algorithm as a set of syntax-directed rules in
\cref{rules:kinding}.
Similar to the syntax-directed rules, higher-kinds are not generally supported
and can only appear by looking-up the kind scheme of a type constructor,
for use in the type application rule {\sc KApp}.
Type variables must be of a simple kind in rule {\sc KVar}.
Kind schemes are instantiated in the {\sc KVar} rules by creating
fresh kind variables and the associated substitution.
{\sc KArr} and {\sc KBorrow} simply returns the kind of the primitive
arrow and borrow types.
